---
title: "Introduction to Statisfy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Statisfy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = T, 
                      comment = ">")
options(tibble.print_min = 4L, tibble.print_max = 4L)
```

```{r}
library(Statisfy)
```
Welcome to the **Statisfy** (A statistician's best friend) package vignette! This document introduces the 
functionalities of the package and demonstrates how its functions can streamline 
common statistical tasks.

To showcase the utility of our functions, we will use the dataset `thesis_data`. 
This dataset contains real data on the environmental conditions of a variety of 
ecosystems throughout Odzala-Kokoua National Park in the Republic of Congo, 
which is being utilized by Dariana for her undergraduate thesis.

### Data: thesis_data
The `thesis_data` dataset contains a range of values, but for our purposes we'll 
focus on `temp_air` and `rel_humidity`

```{r}
head(thesis_data)
```

### Understand distribution with `checkNormality()`
`checkNormality()` helps you evaluate whether a numeric variable follows a 
normal distribution by taking in a numeric variable from a dataset and using it 
to output a histogram, a QQ plot, and a Shapiro-Wilk test result. These are 
typically all of the normality tests we have to run before performing statistical 
analysis, so this function is designed to streamline this precursive step.

In the following example, we utilize `checkNormality()` to visualize air 
temperature follows a normal distribution. 

NOTE TO GRADERS: SINCE MOST OF THESE FUNCTIONS REQUIRE THE USE OF NUMERIC DATA,
IT IS POSSIBLE THAT RUNNING THEM WITH THE MOUSE DATA WILL YIELD AN ERROR MESSAGE
(BUILT INTO THE FUNCTIONS).

```{r}
checkNormality(data = thesis_data, temp_air, line_color = "red")
```

With the help of the variety of graphs and results output, we can see that air
temperature in fact does not follow a normal distribution.

### Choose Appropriate Tests with `whichTest()`
`whichTest()` helps you select the correct statistical test to use for data 
analysis based on the type of data the variable is, the levels of the variable
that exist, and the number of variables you wish to analyze. The function will 
output two of the following recommendations (one parametric test and one 
non-parametric test):

* ANOVA
* Bar/Pie chart
* Chi-Square test
* Independent t-test
* Kruskal-Wallis Test
* Linear Regression
* Mann-Whitney Test
* Paired t-test
* Pearson Correlation Test
* Shapiro-Wilk Test
* Spearman Correlation Test
* U-test
* Wilcoxon Signed-Ranked Test

If we were interested in the variables `temp_air` and `ecosystem` to analyze
whether a relationship exists between air temperature and type of ecosystem but 
were unsure of which test to use, we would use `whichTest()` and receive an
immediate and reliable recommendation.

```{r}
whichTest(data = thesis_data, var1 = "temp_air", var2 = "ecosystem")
```
This recommendation is correct since ecosystem type is a categorical variable
with 3 levels while air temperature is a numeric variable. 

In the case that the function cannot make an appropriate recommendation, the following message will be displayed:

```{r}
## Error in whichTest(df, var1, var2) : This variable type is not supported. Please check your variable type for errors and try again.
```

### Identify outliers with `outliers()`
`outliers()` is an easy way to identify outliers within a dataset based on desired thresholds. By default, the function will use 99th percentile and the 1st percentile when determining what to classify as an outlier.

```{r}
# Outliers that are in the above the 99th percentile and below the 1st percentile
outliers(thesis_data$temp_air, upper = 0.95, lower = 0.05)
```

The user may set their desired threshold to classify an observation as an outlier by changing the function in the following way

```{r}
# Outliers that are above the 89th percentile and below the 12th percentile
outliers(thesis_data$temp_air, 0.89, 0.12)
```

### Summary statistics with `stats()`
`stats()` provides a detailed list of summary statistics of a variable, given that it is numeric. These summary statistics are designed to make it easier to choose between more complicated forms of statistical analysis that require knowledge of data dispersion (ex: beta regression). The function will provide the following statistics:

* Mean
* Variance
* Standard Deviation
* Minimum value
* Maximum value
* 25th percentile
* Median
* 75th percentile
* Coefficient of Variation
 
```{r}
stats(thesis_data$rel_humidity)
```
